---
title: "RI-CLPM & Extensions using Lavaan"
output: 
  html_document:
    toc:  true
    toc_float: true
    code_folding: show
bibliography: references.bib
---

---

## Setup
Below you can find the code for installing and loading the required package `Lavaan` [@lavaan], as well as for reading in the data for the Random Intercept Cross-Lagged Panel Model (RI-CLPM) and its 3 extensions. You can specify the path to the data yourself, or through a menu by using the `file.choose()`-function. You can download the simulated example datasets [here](https://jeroendmulder.github.io/RICLPM/data). 

```{r setup, eval = F, message = F}
# If necessary, install the 'Lavaan' package. 
# install.packages('lavaan', dependencies = T)

# Load the required packages. 
require(lavaan) 

# Load in the data. 
## Traditional RI-CLPM
dat <- read.table("C:/Users/5879167/surfdrive/Research/RI-CLPM Teacher Corner/data/RICLPM.dat", 
                  col.names = c("x1", "y1", "x2", "y2", "x3", "y3", "x4", "y4", "x5", "y5")) 

# Extension 1
datZ <- read.table("C:/Users/5879167/surfdrive/Research/RI-CLPM Teacher Corner/data/RICLPM-Z.dat",
                   col.names = c("x1", "y1", "x2", "y2", "x3", "y3", "x4", "y4", "x5", "y5", "z1", "z2"))

## Extension 2
datMG <- read.table("C:/Users/5879167/surfdrive/Research/RI-CLPM Teacher Corner/data/RICLPM-MG.dat",
                    col.names = c("x1", "y1", "x2", "y2", "x3", "y3", "x4", "y4", "x5", "y5", "G")) 


# Extension 3
datMI <- read.table("C:/Users/5879167/surfdrive/Research/RI-CLPM Teacher Corner/data/RICLPM-MI.dat",
                   col.names = c("x11", "x12", "x13",
                                 "x21", "x22", "x23",
                                 "x31", "x32", "x33",
                                 "x41", "x42", "x43",
                                 "x51", "x52", "x53", 
                                
                                 "y11", "y12", "y13",
                                 "y21", "y22", "y23",
                                 "y31", "y32", "y33", 
                                 "y41", "y42", "y43", 
                                 "y51", "y52", "y53"))

```

---

## The RI-CLPM {.tabset .tabset-fade}
To specify the RI-CLPM we need four parts. 

1. A *between part*, consisting of the random intercepts. It is specified  using the `=~` command, `RIx =~ 1*x1 1*x2 ...`, where `1*` fixes the factor loading to one.
2. A *within part*, consisting of within-unit fluctuations. It is also specified using the `=~` command, `wx1 =~ 1*x1; wx2 =~ 1*x2; ...`. 
3. The *lagged regressions* between the within-unit components, using `wx2 ~ wx1 wy1; wx3 ~ wx2 wy2; ...`. 
4. Relevant *covariances* in both the between and within part. In the within part the components at wave 1, and their residuals at waves 2 and further are correlated within each wave, using `wx1 ~~ wy1; wx2 ~~ wy2;...`. We also need to specify their (residual) variances here using `wx1 ~~ wx1; wx2 ~~ wx2; ...`. For the between part we have to specify the variances and covariance of the random intercepts using `RIx ~~ RIy;`. 

### The basic model
The code for specifying the basic RI-CLPM is given below. 

```{r RICLPM, eval = F}
RICLPM <- '
  # Create two individual factors (random intercepts)
  RIx =~ 1*x1 + 1*x2 + 1*x3 + 1*x4 + 1*x5
  RIy =~ 1*y1 + 1*y2 + 1*y3 + 1*y4 + 1*y5
  
  # Create within-person centered variables
  wx1 =~ 1*x1
  wx2 =~ 1*x2
  wx3 =~ 1*x3 
  wx4 =~ 1*x4
  wx5 =~ 1*x5

  wy1 =~ 1*y1
  wy2 =~ 1*y2
  wy3 =~ 1*y3
  wy4 =~ 1*y4
  wy5 =~ 1*y5

  # Estimate the lagged effects between the within-person centered variables.
  wx2 + wy2 ~ wx1 + wy1
  wx3 + wy3 ~ wx2 + wy2
  wx4 + wy4 ~ wx3 + wy3
  wx5 + wy5 ~ wx4 + wy4

  # Estimate the covariance between the within-person centered variables at the first wave. 
  wx1 ~~ wy1 # Covariance
  
  # Estimate the covariances between the residuals of the within-person centered variables (the innovations).
  wx2 ~~ wy2
  wx3 ~~ wy3
  wx4 ~~ wy4
  wx5 ~~ wy5
  
  # Estimate the variance and covariance of the individual factors. 
  RIx ~~ RIx
  RIy ~~ RIy
  RIx ~~ RIy

  # Estimate the (residual) variance of the within-person centered variables.
  wx1 ~~ wx1 # Variances
  wy1 ~~ wy1 
  wx2 ~~ wx2 # Residual variances
  wy2 ~~ wy2 
  wx3 ~~ wx3 
  wy3 ~~ wy3 
  wx4 ~~ wx4 
  wy4 ~~ wy4 
  wx5 ~~ wx5
  wy5 ~~ wy5
'

RICLPM.fit <- lavaan(RICLPM, data = dat, missing = 'ML',
                     meanstructure = T, int.ov.free = T) 

summary(RICLPM.fit, standardized = T)
```

---

### Constraints over time
Imposing constraints to the model can be achieved through **premultiplication**. It means that we have to add the number that we want to fix the parameter to and an asterisk, to the parameter in the model specification. For example, if we want to fix the factor loading of item `x1` to factor `F` to 0, we need to add `0*` to `F =~ x1`, resulting in `F =~ 0*x1`. Using premultiplication we can also constrain parameters to be the same by giving them the same label. Below we specify a RI-CLPM with the following constraints: 

1) fixed auto-regressive and cross-lagged relations over time, `wx2 ~ a*wx1 + b*wy1; ...`, 
2) time-invariant (residual) (co-)variances in the within-person part `wx2 ~~ cov*wy2; ...`, `wx2 ~~ vx*wx2; ...`, and `wy2 ~~ vy*wy2; ...`, and 
3) constrained grand means over time, `x1 + ... ~ mx*1` and `y1 + ... ~ my*1`. 


```{r RICLPM1, eval = F, echo = F}

# Unconstrained factor loadings for the random intercept. 

RICLPM1 <- '
  # Create two individual factors (random intercepts)
  RIx =~ 1*x1 + x2 + x3 + x4 + x5
  RIy =~ 1*y1 + y2 + y3 + y4 + y5
  
  # Create within-person centered variables
  wx1 =~ 1*x1
  wx2 =~ 1*x2
  wx3 =~ 1*x3 
  wx4 =~ 1*x4
  wx5 =~ 1*x5

  wy1 =~ 1*y1
  wy2 =~ 1*y2
  wy3 =~ 1*y3
  wy4 =~ 1*y4
  wy5 =~ 1*y5

  # Estimate the lagged effects between the within-person centered variables.
  wx2 + wy2 ~ wx1 + wy1
  wx3 + wy3 ~ wx2 + wy2
  wx4 + wy4 ~ wx3 + wy3
  wx5 + wy5 ~ wx4 + wy4

  # Estimate the covariances between the residuals of the within-person centered variables (the innovations).
  wx2 ~~ wy2
  wx3 ~~ wy3
  wx4 ~~ wy4 
  wx5 ~~ wy5

  # Estimate the covariance between the within-person centered variables at the first wave. 
  wx1 ~~ wy1 # Covariance
  
  # Estimate the variance and covariance of the individual factors. 
  RIx ~~ RIx
  RIy ~~ RIy
  RIx ~~ RIy

  # Estimate the (residual) variance of the within-person centered variables.
  wx1 ~~ wx1 # Variance
  wy1 ~~ wy1 
  wx2 ~~ wx2 # Residual variance
  wy2 ~~ wy2 
  wx3 ~~ wx3 
  wy3 ~~ wy3 
  wx4 ~~ wx4 
  wy4 ~~ wy4 
  wx5 ~~ wx5
  wy5 ~~ wy5
'

RICLPM1.fit <- lavaan(RICLPM1, data = dat, missing = 'ML',
                     meanstructure = T, int.ov.free = T) 

summary(RICLPM1.fit, standardized = T)
```

```{r RICLPM2, eval = F, echo = F}

# Fixed autoregressive and cross-lagged relations over time.  

RICLPM2 <- '
  # Create two individual factors (random intercepts)
  RIx =~ 1*x1 + 1*x2 + 1*x3 + 1*x4 + 1*x5
  RIy =~ 1*y1 + 1*y2 + 1*y3 + 1*y4 + 1*y5
  
  # Create within-person centered variables
  wx1 =~ 1*x1
  wx2 =~ 1*x2
  wx3 =~ 1*x3 
  wx4 =~ 1*x4
  wx5 =~ 1*x5

  wy1 =~ 1*y1
  wy2 =~ 1*y2
  wy3 =~ 1*y3
  wy4 =~ 1*y4
  wy5 =~ 1*y5

  # Estimate the lagged effects between the within-person centered variables.
  wx2 ~ a*wx1 + b*wy1 
  wy2 ~ c*wx1 + d*wy1
  wx3 ~ a*wx2 + b*wy2
  wy3 ~ c*wx2 + d*wy2
  wx4 ~ a*wx3 + b*wy3
  wy4 ~ c*wx3 + d*wy3
  wx5 ~ a*wx4 + b*wy4
  wy5 ~ c*wx4 + d*wy4

  # Estimate the covariances between the residuals of the within-person centered variables (the innovations).
  wx2 ~~ wy2
  wx3 ~~ wy3
  wx4 ~~ wy4 
  wx5 ~~ wy5

  # Estimate the covariance between the within-person centered variables at the first wave. 
  wx1 ~~ wy1 # Covariance
  
  # Estimate the variance and covariance of the individual factors. 
  RIx ~~ RIx
  RIy ~~ RIy
  RIx ~~ RIy

  # Estimate the (residual) variance of the within-person centered variables.
  wx1 ~~ wx1 # Variance
  wy1 ~~ wy1 
  wx2 ~~ wx2 # Residual variance
  wy2 ~~ wy2 
  wx3 ~~ wx3 
  wy3 ~~ wy3 
  wx4 ~~ wx4 
  wy4 ~~ wy4 
  wx5 ~~ wx5
  wy5 ~~ wy5

'

RICLPM2.fit <- lavaan(RICLPM2, data = dat, missing = 'ML',
                     meanstructure = T, int.ov.free = T) 

summary(RICLPM2.fit, standardized = T)
```

```{r RICLPM3, eval = F, echo = F}

# Constrained grand means. 

RICLPM3 <- '
  # Create two individual factors (random intercepts)
  RIx =~ 1*x1 + 1*x2 + 1*x3 + 1*x4 + 1*x5
  RIy =~ 1*y1 + 1*y2 + 1*y3 + 1*y4 + 1*y5
  
  # Create within-person centered variables
  wx1 =~ 1*x1
  wx2 =~ 1*x2
  wx3 =~ 1*x3 
  wx4 =~ 1*x4
  wx5 =~ 1*x5

  wy1 =~ 1*y1
  wy2 =~ 1*y2
  wy3 =~ 1*y3
  wy4 =~ 1*y4
  wy5 =~ 1*y5

  # Estimate the lagged effects between the within-person centered variables.
  wx2 ~ wx1 + wy1 
  wy2 ~ wx1 + wy1
  wx3 ~ wx2 + wy2
  wy3 ~ wx2 + wy2
  wx4 ~ wx3 + wy3
  wy4 ~ wx3 + wy3
  wx5 ~ wx4 + wy4
  wy5 ~ wx4 + wy4

  # Estimate the covariances between the residuals of the within-person centered variables (the innovations).
  wx2 ~~ wy2
  wx3 ~~ wy3
  wx4 ~~ wy4 
  wx5 ~~ wy5

  # Estimate the covariance between the within-person centered variables at the first wave. 
  wx1 ~~ wy1 # Covariance
  
  # Estimate the variance and covariance of the individual factors. 
  RIx ~~ RIx
  RIy ~~ RIy
  RIx ~~ RIy

  # Estimate the (residual) variance of the within-person centered variables.
  wx1 ~~ wx1 # Variance
  wy1 ~~ wy1 
  wx2 ~~ wx2 # Residual variance
  wy2 ~~ wy2 
  wx3 ~~ wx3 
  wy3 ~~ wy3 
  wx4 ~~ wx4 
  wy4 ~~ wy4 
  wx5 ~~ wx5
  wy5 ~~ wy5
  
  # Constrain the grand means over time. 
  x1 + x2 + x3 + x4 + x5 ~ mx*1
  y1 + y2 + y3 + y4 + y5 ~ my*1
'

RICLPM3.fit <- lavaan(RICLPM3, data = dat, missing = 'ML',
                      meanstructure = T, int.ov.free = T) 

summary(RICLPM3.fit, standardized = T)
```

```{r RICLPM5, eval = F, echo = T}

RICLPM5 <- '
  # Create two individual factors (random intercepts)
  RIx =~ 1*x1 + 1*x2 + 1*x3 + 1*x4 + 1*x5
  RIy =~ 1*y1 + 1*y2 + 1*y3 + 1*y4 + 1*y5
  
  # Create within-person centered variables
  wx1 =~ 1*x1
  wx2 =~ 1*x2
  wx3 =~ 1*x3 
  wx4 =~ 1*x4
  wx5 =~ 1*x5

  wy1 =~ 1*y1
  wy2 =~ 1*y2
  wy3 =~ 1*y3
  wy4 =~ 1*y4
  wy5 =~ 1*y5

  # Estimate the lagged effects between the within-person centered variables (constrained).
  wx2 ~ a*wx1 + b*wy1 
  wy2 ~ c*wx1 + d*wy1
  wx3 ~ a*wx2 + b*wy2
  wy3 ~ c*wx2 + d*wy2
  wx4 ~ a*wx3 + b*wy3
  wy4 ~ c*wx3 + d*wy3
  wx5 ~ a*wx4 + b*wy4
  wy5 ~ c*wx4 + d*wy4

  # Estimate the covariances between the residuals of the within-person centered variables 
  # (the innovations, constrained).
  wx2 ~~ cov*wy2
  wx3 ~~ cov*wy3
  wx4 ~~ cov*wy4 
  wx5 ~~ cov*wy5

  # Estimate the covariance between the within-person centered variables at the first wave. 
  wx1 ~~ wy1 # Covariance
  
  # Estimate the variance and covariance of the individual factors. 
  RIx ~~ RIx
  RIy ~~ RIy
  RIx ~~ RIy

  # Estimate the (residual) variance of the within-person centered variables (constrained).
  wx1 ~~ wx1 # Variance
  wy1 ~~ wy1 
  wx2 ~~ vx*wx2 # Residual variance
  wy2 ~~ vy*wy2 
  wx3 ~~ vx*wx3 
  wy3 ~~ vy*wy3 
  wx4 ~~ vx*wx4 
  wy4 ~~ vy*wy4 
  wx5 ~~ vx*wx5
  wy5 ~~ vy*wy5
  
  # Constrain the grand means over time. 
  x1 + x2 + x3 + x4 + x5 ~ mx*1
  y1 + y2 + y3 + y4 + y5 ~ my*1
'

RICLPM5.fit <- lavaan(RICLPM5, data = dat, missing = 'ML',
                      meanstructure = T, int.ov.free = T) 

summary(RICLPM5.fit, standardized = T)
```

---

## Ext. 1: time-invariant predictor {.tabset .tabset-fade}
Use the tabs below to navigate to the model specification of the RI-CLPM with

- a time-invariant predictor $z_{1}$ of the observed variables,
- a time-invariant predictor $z_{1}$ of the random intercepts, 
- a time-invariant outcome $z_{2}$ of the observed variables, or
- a time-invariant outcome $z_{2}$ of the random intercepts. 

You can download the simulated example dataset *RICLPMZ.dat* [here](https://jeroendmulder.github.io/RI-CLPM/data/RICLPMZ.dat). 

### $z_{1}$ predicting observed
Below you can find the code for a RI-CLPM with 5 waves and a time-invariant predictor $z_{1}$ for the observed variables. The effect of $z_{1}$ on the observed variables is constrained to be the same across waves.

```{r RICLPM-Z, eval = F}
RICLPM.ext1 <- '
  # Create two individual factors (random intercepts)
  RIx =~ 1*x1 + 1*x2 + 1*x3 + 1*x4 + 1*x5
  RIy =~ 1*y1 + 1*y2 + 1*y3 + 1*y4 + 1*y5
  
  # Create within-person centered variables
  wx1 =~ 1*x1
  wx2 =~ 1*x2
  wx3 =~ 1*x3 
  wx4 =~ 1*x4
  wx5 =~ 1*x5

  wy1 =~ 1*y1
  wy2 =~ 1*y2
  wy3 =~ 1*y3
  wy4 =~ 1*y4
  wy5 =~ 1*y5

  # Regression of observed variables on z1 (constrained). 
  x1 + x2 + x3 + x4 + x5 ~ s1*z1 # Constrained over time. 
  y1 + y2 + y3 + y4 + y5 ~ s2*z1 # Constrained over time. 

  # Estimate the lagged effects between the within-person centered variables.
  wx2 + wy2 ~ wx1 + wy1
  wx3 + wy3 ~ wx2 + wy2
  wx4 + wy4 ~ wx3 + wy3
  wx5 + wy5 ~ wx4 + wy4

  # Estimate the covariance between the within-person centered variables at the first wave. 
  wx1 ~~ wy1 # Covariance
  
  # Estimate the covariances between the residuals of the within-person centered variables (the innovations).
  wx2 ~~ wy2
  wx3 ~~ wy3
  wx4 ~~ wy4
  wx5 ~~ wy5
  
  # Estimate the variance and covariance of the individual factors. 
  RIx ~~ RIx
  RIy ~~ RIy
  RIx ~~ RIy

  # Estimate the (residual) variance of the within-person centered variables.
  wx1 ~~ wx1 # Variances
  wy1 ~~ wy1 
  wx2 ~~ wx2 # Residual variances
  wy2 ~~ wy2 
  wx3 ~~ wx3 
  wy3 ~~ wy3 
  wx4 ~~ wx4 
  wy4 ~~ wy4 
  wx5 ~~ wx5
  wy5 ~~ wy5
'

RICLPM.ext1.fit <- lavaan(RICLPM.ext1, data = datZ, missing = 'ML',
                          meanstructure = T, int.ov.free = T) 

summary(RICLPM.ext1.fit, standardized = T)
```

---

### $z_{1}$ predicting RIs
Below you can find the code for a RI-CLPM with 5 waves and a time-invariant predictor $z_{1}$ for the random intercepts. 
```{r RICLPM-Z1, eval = F}

RICLPM1.ext1 <- '
  # Create two individual factors (random intercepts)
  RIx =~ 1*x1 + 1*x2 + 1*x3 + 1*x4 + 1*x5
  RIy =~ 1*y1 + 1*y2 + 1*y3 + 1*y4 + 1*y5
  
  # Create within-person centered variables
  wx1 =~ 1*x1
  wx2 =~ 1*x2
  wx3 =~ 1*x3 
  wx4 =~ 1*x4
  wx5 =~ 1*x5

  wy1 =~ 1*y1
  wy2 =~ 1*y2
  wy3 =~ 1*y3
  wy4 =~ 1*y4
  wy5 =~ 1*y5

  # Regression of random intercepts on z1. 
  RIx ~ z1 
  RIy ~ z1 

  # Estimate the lagged effects between the within-person centered variables.
  wx2 + wy2 ~ wx1 + wy1
  wx3 + wy3 ~ wx2 + wy2
  wx4 + wy4 ~ wx3 + wy3
  wx5 + wy5 ~ wx4 + wy4

  # Estimate the covariance between the within-person centered variables at the first wave. 
  wx1 ~~ wy1 # Covariance
  
  # Estimate the covariances between the residuals of the within-person centered variables (the innovations).
  wx2 ~~ wy2
  wx3 ~~ wy3
  wx4 ~~ wy4
  wx5 ~~ wy5
  
  # Estimate the variance and covariance of the individual factors. 
  RIx ~~ RIx
  RIy ~~ RIy
  RIx ~~ RIy

  # Estimate the (residual) variance of the within-person centered variables.
  wx1 ~~ wx1 # Variances
  wy1 ~~ wy1 
  wx2 ~~ wx2 # Residual variances
  wy2 ~~ wy2 
  wx3 ~~ wx3 
  wy3 ~~ wy3 
  wx4 ~~ wx4 
  wy4 ~~ wy4 
  wx5 ~~ wx5
  wy5 ~~ wy5
'

RICLPM1.ext1.fit <- lavaan(RICLPM1.ext1, data = datZ, missing = 'ML', 
                           meanstructure = T, int.ov.free = T) 

summary(RICLPM1.ext1.fit, standardized = T)
```

---

### observed predicting $z_{2}$
Below you can find the code for a RI-CLPM with 5 waves and the observed variables predicting the time-invariant outcome $z_{2}$. The effect of the observed variables on the outcome $z_{2}$ is constrained to be the same across waves. 

```{r RICLPM-Z2, eval = F}
RICLPM2.ext1 <- '
  # Create two individual factors (random intercepts)
  RIx =~ 1*x1 + 1*x2 + 1*x3 + 1*x4 + 1*x5
  RIy =~ 1*y1 + 1*y2 + 1*y3 + 1*y4 + 1*y5
  
  # Create within-person centered variables
  wx1 =~ 1*x1
  wx2 =~ 1*x2
  wx3 =~ 1*x3 
  wx4 =~ 1*x4
  wx5 =~ 1*x5

  wy1 =~ 1*y1
  wy2 =~ 1*y2
  wy3 =~ 1*y3
  wy4 =~ 1*y4
  wy5 =~ 1*y5

  # Regression of z2 on observed (constrained). 
  z2 ~ s1*x1 + s1*x2 + s1*x3 + s1*x4 + s1*x5 # Constrained over time. 
  z2 ~ s2*y1 + s2*y2 + s2*y3 + s2*y4 + s2*y5 # Constrained over time. 

  # Estimate the lagged effects between the within-person centered variables.
  wx2 + wy2 ~ wx1 + wy1
  wx3 + wy3 ~ wx2 + wy2
  wx4 + wy4 ~ wx3 + wy3
  wx5 + wy5 ~ wx4 + wy4

  # Estimate the covariance between the within-person centered variables at the first wave. 
  wx1 ~~ wy1 # Covariance
  
  # Estimate the covariances between the residuals of the within-person centered variables (the innovations).
  wx2 ~~ wy2
  wx3 ~~ wy3
  wx4 ~~ wy4
  wx5 ~~ wy5
  
  # Estimate the variance and covariance of the individual factors. 
  RIx ~~ RIx
  RIy ~~ RIy
  RIx ~~ RIy

  # Estimate the (residual) variance of the within-person centered variables.
  wx1 ~~ wx1 # Variances
  wy1 ~~ wy1 
  wx2 ~~ wx2 # Residual variances
  wy2 ~~ wy2 
  wx3 ~~ wx3 
  wy3 ~~ wy3 
  wx4 ~~ wx4 
  wy4 ~~ wy4 
  wx5 ~~ wx5
  wy5 ~~ wy5
'

RICLPM2.ext1.fit <- lavaan(RICLPM2.ext1, data = datZ, missing = 'ML',
                          meanstructure = T, int.ov.free = T) 

summary(RICLPM2.ext1.fit, standardized = T)
```

### RIs predicting $z_{2}$
Below you can find the code for a RI-CLPM with 5 waves and the random intercepts predicting the time-invariant outcome $z_{2}$.

```{r RICLPM-Z3, eval = F}
RICLPM3.ext1 <- '
  # Create two individual factors (random intercepts)
  RIx =~ 1*x1 + 1*x2 + 1*x3 + 1*x4 + 1*x5
  RIy =~ 1*y1 + 1*y2 + 1*y3 + 1*y4 + 1*y5
  
  # Create within-person centered variables
  wx1 =~ 1*x1
  wx2 =~ 1*x2
  wx3 =~ 1*x3 
  wx4 =~ 1*x4
  wx5 =~ 1*x5

  wy1 =~ 1*y1
  wy2 =~ 1*y2
  wy3 =~ 1*y3
  wy4 =~ 1*y4
  wy5 =~ 1*y5

  # Regression of z2 on random intercepts (constrained). 
  z2 ~ RIx + RIy
  
  # Estimate the lagged effects between the within-person centered variables.
  wx2 + wy2 ~ wx1 + wy1
  wx3 + wy3 ~ wx2 + wy2
  wx4 + wy4 ~ wx3 + wy3
  wx5 + wy5 ~ wx4 + wy4

  # Estimate the covariance between the within-person centered variables at the first wave. 
  wx1 ~~ wy1 # Covariance
  
  # Estimate the covariances between the residuals of the within-person centered variables (the innovations).
  wx2 ~~ wy2
  wx3 ~~ wy3
  wx4 ~~ wy4
  wx5 ~~ wy5
  
  # Estimate the variance and covariance of the individual factors. 
  RIx ~~ RIx
  RIy ~~ RIy
  RIx ~~ RIy

  # Estimate the (residual) variance of the within-person centered variables.
  wx1 ~~ wx1 # Variances
  wy1 ~~ wy1 
  wx2 ~~ wx2 # Residual variances
  wy2 ~~ wy2 
  wx3 ~~ wx3 
  wy3 ~~ wy3 
  wx4 ~~ wx4 
  wy4 ~~ wy4 
  wx5 ~~ wx5
  wy5 ~~ wy5
'

RICLPM3.ext1.fit <- lavaan(RICLPM3.ext1, data = datZ, missing = 'ML',
                          meanstructure = T, int.ov.free = T) 

summary(RICLPM3.ext1.fit, standardized = T)
```

## Ext. 2: multiple group {.tabset .tabset-fade}
Use the tabs below to navigate to the model specification of the basic multiple-group model, or the model with constrained lagged parameters (and intercepts across groups). You can download the simulated example dataset *MGRICLPM.dat* [here](https://jeroendmulder.github.io/RICLPM/data/MGRICLPM.dat).

### The basic model
Below you can find the code for a multiple group RI-CLPM with 5 waves.

```{r RICLPM-MG, eval = F}
RICLPM.ext2 <- '
  # Create two individual factors (random intercepts)
  RIx =~ 1*x1 + 1*x2 + 1*x3 + 1*x4 + 1*x5
  RIy =~ 1*y1 + 1*y2 + 1*y3 + 1*y4 + 1*y5
  
  # Create within-person centered variables
  wx1 =~ 1*x1
  wx2 =~ 1*x2
  wx3 =~ 1*x3 
  wx4 =~ 1*x4
  wx5 =~ 1*x5

  wy1 =~ 1*y1
  wy2 =~ 1*y2
  wy3 =~ 1*y3
  wy4 =~ 1*y4
  wy5 =~ 1*y5

  # Estimate the lagged effects between the within-person centered variables.
  wx2 + wy2 ~ wx1 + wy1
  wx3 + wy3 ~ wx2 + wy2
  wx4 + wy4 ~ wx3 + wy3
  wx5 + wy5 ~ wx4 + wy4

  # Estimate the covariance between the within-person centered variables at the first wave. 
  wx1 ~~ wy1 # Covariance
  
  # Estimate the covariances between the residuals of the within-person centered variables (the innovations).
  wx2 ~~ wy2
  wx3 ~~ wy3
  wx4 ~~ wy4
  wx5 ~~ wy5
  
  # Estimate the variance and covariance of the individual factors. 
  RIx ~~ RIx
  RIy ~~ RIy
  RIx ~~ RIy

  # Estimate the (residual) variance of the within-person centered variables.
  wx1 ~~ wx1 # Variances
  wy1 ~~ wy1 
  wx2 ~~ wx2 # Residual variances
  wy2 ~~ wy2 
  wx3 ~~ wx3 
  wy3 ~~ wy3 
  wx4 ~~ wx4 
  wy4 ~~ wy4 
  wx5 ~~ wx5
  wy5 ~~ wy5
'

RICLPM.ext2.fit <- lavaan(RICLPM.ext2, data = datMG, missing = 'ML',
                           group = "G", 
                           meanstructure = T, int.ov.free = T)
summary(RICLPM.ext2.fit)

```

---

### Constrained lagged-parameters
Below you can find the code for a multiple group RI-CLPM with 4 waves. The lagged-parameters are constrained to be equal over time.

```{r RICLPM-MG1, eval = F}
RICLPM1.ext2 <- '
  # Create two individual factors (random intercepts)
  RIx =~ 1*x1 + 1*x2 + 1*x3 + 1*x4 + 1*x5
  RIy =~ 1*y1 + 1*y2 + 1*y3 + 1*y4 + 1*y5
  
  # Create within-person centered variables
  wx1 =~ 1*x1
  wx2 =~ 1*x2
  wx3 =~ 1*x3 
  wx4 =~ 1*x4
  wx5 =~ 1*x5

  wy1 =~ 1*y1
  wy2 =~ 1*y2
  wy3 =~ 1*y3
  wy4 =~ 1*y4
  wy5 =~ 1*y5

  # Estimate the lagged effects between the within-person centered variables. Constrain the  
  # autoregressive effects across groups. 
  wx2 ~ c(a1, a1)*wx1 + c(b1, b1)*wy1
  wy2 ~ c(c1, c1)*wx1 + c(d1, d1)*wy1
  wx3 ~ c(a2, a2)*wx2 + c(b2, b2)*wy2
  wy3 ~ c(c2, c2)*wx2 + c(d2, d2)*wy2
  wx4 ~ c(a3, a3)*wx3 + c(b3, b3)*wy3
  wy4 ~ c(c3, c3)*wx3 + c(d3, d3)*wy3 
  wx5 ~ c(a4, a4)*wx4 + c(b4, b4)*wy4
  wy5 ~ c(c4, c4)*wx4 + c(d4, d4)*wy4

  # Estimate the covariance between the within-person centered variables at the first wave. 
  wx1 ~~ wy1 # Covariance
  
  # Estimate the covariances between the residuals of the within-person centered variables (the innovations).
  wx2 ~~ wy2
  wx3 ~~ wy3
  wx4 ~~ wy4
  wx5 ~~ wy5
  
  # Estimate the variance and covariance of the individual factors. 
  RIx ~~ RIx
  RIy ~~ RIy
  RIx ~~ RIy

  # Estimate the (residual) variance of the within-person centered variables.
  wx1 ~~ wx1 # Variances
  wy1 ~~ wy1 
  wx2 ~~ wx2 # Residual variances
  wy2 ~~ wy2 
  wx3 ~~ wx3 
  wy3 ~~ wy3 
  wx4 ~~ wx4 
  wy4 ~~ wy4 
  wx5 ~~ wx5
  wy5 ~~ wy5
'

RICLPM1.ext2.fit <- lavaan(RICLPM1.ext2, data = datMG, missing = 'ML',
                           group = "G", 
                           meanstructure = T, int.ov.free = T)
summary(RICLPM1.ext2.fit)

```

---

## Ext. 3: multiple indicator {.tabset .tabset-fade}
Use the tabs below to navigate to the model specification of a multiple indicator RI-CLPM, 5 waves and 3 indicators for each variable at each wave. The five steps correspond to:

- the initial model (Step 1),
- weak factorial invariance (Step 2),
- strong factorial invariance (Step 3),
- strong factorial invariance with factor loadings equal to the within-person factor loadings (Extra), and
- the latent RI-CLPM (Step 4).

You can download the simulated example dataset *MIRICLPM.dat* [here](https://jeroendmulder.github.io/RI-CLPM/data/MIRICLPM.dat).

### Step 1
When we have three indicators $X$, measured at five waves, we specify three random intercepts to capture the trait-like part of each indicator, that is, `RIX1 =~ 1*x11 1*x21 ...`, `RIX2 =~ 1*x121 1*x22@1 ...`, and `RIX3 =~ 1*x13 1*x23 ...`. In addition, we specify five within-unit components that capture the state-like part at each wave, using `WFX1 =~ x11 x12 x13; WFX2 =~ x21 x22 x23; ...`.

At the latent within-unit level, we specify the dynamic model in Mplus using `WFX2 ~ WFY1 + WFX1; WFX3 ~ WFY2 + WFX2; ...`. In addition, we allow the within-person factors at the first wave, and their residuals at subsequent waves to be correlated within each wave, `WFX1 ~~ WFY1; WFX2 ~~ WFY2; ...`. The six random intercepts are allowed to be freely correlated with each other through `RIX1 + RIX2 + ... ~~ RIX1 + RIX2 + ...`. 

```{r RICLPM-MI1, eval = F, warning = F}
RICLPM1.ext3 <- '
  
  ################
  # BETWEEN PART #
  ################
  
  # Create between factors (random intercepts) for each indicator separately.
  RIX1 =~ 1*x11 + 1*x21 + 1*x31 + 1*x41 + 1*x51
  RIX2 =~ 1*x12 + 1*x22 + 1*x32 + 1*x42 + 1*x52
  RIX3 =~ 1*x13 + 1*x23 + 1*x33 + 1*x43 + 1*x53
  
  RIY1 =~ 1*y11 + 1*y21 + 1*y31 + 1*y41 + 1*y51
  RIY2 =~ 1*y12 + 1*y22 + 1*y32 + 1*y42 + 1*y52
  RIY3 =~ 1*y13 + 1*y23 + 1*y33 + 1*y43 + 1*y53
  
  ##################################
  # WITHIN PART: MEASUREMENT MODEL #
  ##################################
  
  # Factor models for x at 5 waves.
  WFX1 =~ x11 + x12 + x13
  WFX2 =~ x21 + x22 + x23
  WFX3 =~ x31 + x32 + x33
  WFX4 =~ x41 + x42 + x43
  WFX5 =~ x51 + x52 + x53
  
  # Factor models for y at 5 waves.
  WFY1 =~ y11 + y12 + y13
  WFY2 =~ y21 + y22 + y23
  WFY3 =~ y31 + y32 + y33
  WFY4 =~ y41 + y42 + y43
  WFY5 =~ y51 + y52 + y53
  
  #########################
  # WITHIN PART: DYNAMICS #
  #########################
  
  # Specify the lagged effects between the within-person centered latent variables.
  WFX2 + WFY2 ~ WFX1 + WFY1
  WFX3 + WFY3 ~ WFX2 + WFY2
  WFX4 + WFY4 ~ WFX3 + WFY3
  WFX5 + WFY5 ~ WFX4 + WFY4
  
  # Estimate the correlations within the same wave.
  WFX1 ~~ WFY1
  WFX2 ~~ WFY2
  WFX3 ~~ WFY3 
  WFX4 ~~ WFY4
  WFX5 ~~ WFY5
  
  ##########################
  # ADDITIONAL CONSTRAINTS #
  ##########################
  
  # Constrain covariance of the between factors and exogenous within factors to 0. 
  RIX1 + RIX2 + RIX3 + RIY1 + RIY2 + RIY3 ~~ 0*WFY1 + 0*WFX1

'
RICLPM1.ext3.fit <- cfa(RICLPM1.ext3, data = datMI, missing = 'ML')

summary(RICLPM1.ext3.fit, standardized = T)
```

---

### Step 2
In second step, we constrain the factor loadings to be invariant over time using the labels `a*`, `b*`, etc.

```{r RICLPM-MI2, eval = F, warning = F}
RICLPM2.ext3 <- '
  
  ################
  # BETWEEN PART #
  ################
  
  # Create between factors (random intercepts) for each indicator separately.
  RIX1 =~ 1*x11 + 1*x21 + 1*x31 + 1*x41 + 1*x51
  RIX2 =~ 1*x12 + 1*x22 + 1*x32 + 1*x42 + 1*x52
  RIX3 =~ 1*x13 + 1*x23 + 1*x33 + 1*x43 + 1*x53
  
  RIY1 =~ 1*y11 + 1*y21 + 1*y31 + 1*y41 + 1*y51
  RIY2 =~ 1*y12 + 1*y22 + 1*y32 + 1*y42 + 1*y52
  RIY3 =~ 1*y13 + 1*y23 + 1*y33 + 1*y43 + 1*y53
  
  ##################################
  # WITHIN PART: MEASUREMENT MODEL #
  ##################################
  
  # Factor models for x at 5 waves (constrained).
  WFX1 =~ a*x11 + b*x12 + c*x13
  WFX2 =~ a*x21 + b*x22 + c*x23
  WFX3 =~ a*x31 + b*x32 + c*x33
  WFX4 =~ a*x41 + b*x42 + c*x43
  WFX5 =~ a*x51 + b*x52 + c*x53
  
  # Factor models for y at 5 waves (constrained).
  WFY1 =~ d*y11 + e*y12 + f*y13
  WFY2 =~ d*y21 + e*y22 + f*y23
  WFY3 =~ d*y31 + e*y32 + f*y33
  WFY4 =~ d*y41 + e*y42 + f*y43
  WFY5 =~ d*y51 + e*y52 + f*y53
  
  #########################
  # WITHIN PART: DYNAMICS #
  #########################
  
  # Specify the lagged effects between the within-person centered latent variables.
  WFX2 + WFY2 ~ WFX1 + WFY1
  WFX3 + WFY3 ~ WFX2 + WFY2
  WFX4 + WFY4 ~ WFX3 + WFY3
  WFX5 + WFY5 ~ WFX4 + WFY4
  
  # Estimate the correlations within the same wave.
  WFX1 ~~ WFY1
  WFX2 ~~ WFY2
  WFX3 ~~ WFY3 
  WFX4 ~~ WFY4
  WFX5 ~~ WFY5
  
  ##########################
  # ADDITIONAL CONSTRAINTS #
  ##########################
  
  # Constrain covariance of the between factors and exogenous within factors to 0. 
  RIX1 + RIX2 + RIX3 + RIY1 + RIY2 + RIY3 ~~ 0*WFY1 + 0*WFX1

'
RICLPM2.ext3.fit <- cfa(RICLPM2.ext3, data = datMI, missing = 'ML')

summary(RICLPM2.ext3.fit, standardized = T)
```

---

### Step 3
Multiple indicator RI-CLPM 4 waves with 3 indicators for each variable at each wave (24 observed variables). Fitting a model with constraints to ensure strong factorial invariance, with a random intercept for each indicator separately.

```{r RICLPM-MI3, eval = F, warning = F}
RICLPM3.ext3 <- '
  
  ################
  # BETWEEN PART #
  ################
  
  # Create between factors (random intercepts) for each indicator separately.
  RIX1 =~ 1*x11 + 1*x21 + 1*x31 + 1*x41 + 1*x51
  RIX2 =~ 1*x12 + 1*x22 + 1*x32 + 1*x42 + 1*x52
  RIX3 =~ 1*x13 + 1*x23 + 1*x33 + 1*x43 + 1*x53
  
  RIY1 =~ 1*y11 + 1*y21 + 1*y31 + 1*y41 + 1*y51
  RIY2 =~ 1*y12 + 1*y22 + 1*y32 + 1*y42 + 1*y52
  RIY3 =~ 1*y13 + 1*y23 + 1*y33 + 1*y43 + 1*y53
  
  ##################################
  # WITHIN PART: MEASUREMENT MODEL #
  ##################################
  
  # Factor models for x at 5 waves (constrained).
  WFX1 =~ a*x11 + b*x12 + c*x13
  WFX2 =~ a*x21 + b*x22 + c*x23
  WFX3 =~ a*x31 + b*x32 + c*x33
  WFX4 =~ a*x41 + b*x42 + c*x43
  WFX5 =~ a*x51 + b*x52 + c*x53
  
  # Factor models for y at 5 waves (constrained).
  WFY1 =~ d*y11 + e*y12 + f*y13
  WFY2 =~ d*y21 + e*y22 + f*y23
  WFY3 =~ d*y31 + e*y32 + f*y33
  WFY4 =~ d*y41 + e*y42 + f*y43
  WFY5 =~ d*y51 + e*y52 + f*y53
  
  # Constrained intercepts over time (this is necessary for strong factorial 
  # invariance; without these contraints we have week factorial invariance). 
  
  x11 + x21 + x31 + x41 + x51 ~ g*1
  x12 + x22 + x32 + x42 + x52 ~ h*1
  x13 + x23 + x33 + x43 + x53 ~ i*1

  y11 + y21 + y31 + y41 + y51 ~ j*1
  y12 + y22 + y32 + y42 + y52 ~ k*1
  y13 + y23 + y33 + y43 + y53 ~ l*1
  
  # Free latent means from t = 2 onward (only do this in combination with the 
  # constraints on the intercepts; without these, this would not be specified).
  
  WFX2 + WFX3 + WFX4 + WFX5 + WFY2 + WFY3 + WFY4 + WFY5 ~ 1
  
  #########################
  # WITHIN PART: DYNAMICS #
  #########################
  
  # Specify the lagged effects between the within-person centered latent variables.
  WFX2 + WFY2 ~ WFX1 + WFY1
  WFX3 + WFY3 ~ WFX2 + WFY2
  WFX4 + WFY4 ~ WFX3 + WFY3
  WFX5 + WFY5 ~ WFX4 + WFY4
  
  # Estimate the correlations within the same wave.
  WFX1 ~~ WFY1
  WFX2 ~~ WFY2
  WFX3 ~~ WFY3 
  WFX4 ~~ WFY4
  WFX5 ~~ WFY5
  
  ##########################
  # ADDITIONAL CONSTRAINTS #
  ##########################
  
  # Constrain covariance of the between factors and exogenous within factors to 0. 
  RIX1 + RIX2 + RIX3 + RIY1 + RIY2 + RIY3 ~~ 0*WFY1 + 0*WFX1

'
RICLPM3.ext3.fit <- cfa(RICLPM3.ext3, data = datMI, missing = 'ML')

summary(RICLPM3.ext3.fit, standardized = T)
```

---

### Extra
Multiple indicator RI-CLPM, 5 waves with 3 indicators for each variable at each wave (30 observed variables). Fitting a model With constraints to ensure strong factorial invariance, with a random intercept for each indicator separately, for which a factor model is specified, with factor loadings equal to the within- person factor loadings. 

```{r RICLPM-MI4, eval = F, warning = F}
RICLPM4.ext3 <- '
  
  ################
  # BETWEEN PART #
  ################
  
  # Create between factors (random intercepts) for each indicator separately.
  RIX1 =~ 1*x11 + 1*x21 + 1*x31 + 1*x41 + 1*x51
  RIX2 =~ 1*x12 + 1*x22 + 1*x32 + 1*x42 + 1*x52
  RIX3 =~ 1*x13 + 1*x23 + 1*x33 + 1*x43 + 1*x53
  
  RIY1 =~ 1*y11 + 1*y21 + 1*y31 + 1*y41 + 1*y51
  RIY2 =~ 1*y12 + 1*y22 + 1*y32 + 1*y42 + 1*y52
  RIY3 =~ 1*y13 + 1*y23 + 1*y33 + 1*y43 + 1*y53
  
  # Create a single random intercept for al x variables, and another for all 
  # y variables and constrain the factor loadings to be identical to the 
  # within-person factor loadings. 
  
  RIX =~ a*RIX1 + b*RIX2 + c*RIX3
  RIY =~ d*RIY1 + e*RIY2 + f*RIY3
  
  # Add (co)variance between two higher-order random intercepts. 
  RIX ~~ RIY
  RIX ~~ RIX
  RIY ~~ RIY
  
  # Constrain measurement error variances of the second order factor model to 0. 
  RIX1 ~~ 0*RIX1
  RIX2 ~~ 0*RIX2
  RIX3 ~~ 0*RIX3

  RIY1 ~~ 0*RIY1
  RIY2 ~~ 0*RIY2
  RIY3 ~~ 0*RIY3
  
  ##################################
  # WITHIN PART: MEASUREMENT MODEL #
  ##################################
  
  # Factor models for x at 5 waves (constrained).
  WFX1 =~ a*x11 + b*x12 + c*x13
  WFX2 =~ a*x21 + b*x22 + c*x23
  WFX3 =~ a*x31 + b*x32 + c*x33
  WFX4 =~ a*x41 + b*x42 + c*x43
  WFX5 =~ a*x51 + b*x52 + c*x53
  
  # Factor models for y at 5 waves (constrained).
  WFY1 =~ d*y11 + e*y12 + f*y13
  WFY2 =~ d*y21 + e*y22 + f*y23
  WFY3 =~ d*y31 + e*y32 + f*y33
  WFY4 =~ d*y41 + e*y42 + f*y43
  WFY5 =~ d*y51 + e*y52 + f*y53
  
  # Constrained intercepts over time (this is necessary for strong factorial 
  # invariance; without these contraints we have week factorial invariance). 
  
  x11 + x21 + x31 + x41 + x51 ~ g*1
  x12 + x22 + x32 + x42 + x52 ~ h*1
  x13 + x23 + x33 + x43 + x53 ~ i*1

  y11 + y21 + y31 + y41 + y51 ~ j*1
  y12 + y22 + y32 + y42 + y52 ~ k*1
  y13 + y23 + y33 + y43 + y53 ~ l*1
  
  # Free latent means from t = 2 onward (only do this in combination with the 
  # constraints on the intercepts; without these, this would not be specified).
  
  WFX2 + WFX3 + WFX4 + WFX5 + WFY2 + WFY3 + WFY4 + WFY5 ~ 1
  
  #########################
  # WITHIN PART: DYNAMICS #
  #########################
  
  # Specify the lagged effects between the within-person centered latent variables.
  WFX2 + WFY2 ~ WFX1 + WFY1
  WFX3 + WFY3 ~ WFX2 + WFY2
  WFX4 + WFY4 ~ WFX3 + WFY3
  WFX5 + WFY5 ~ WFX4 + WFY4
  
  # Estimate the correlations within the same wave.
  WFX1 ~~ WFY1
  WFX2 ~~ WFY2
  WFX3 ~~ WFY3 
  WFX4 ~~ WFY4
  WFX5 ~~ WFY5
  
  ##########################
  # ADDITIONAL CONSTRAINTS #
  ##########################
  
  # Constrain covariance of the between factors and exogenous within factors to 0. 
  RIX + RIY + RIX1 + RIX2 + RIX3 + RIY1 + RIY2 + RIY3 ~~ 0*WFY1 + 0*WFX1

'
RICLPM4.ext3.fit <- cfa(RICLPM4.ext3, data = datMI, missing = 'ML')

summary(RICLPM4.ext3.fit, standardized = T)
```

---

### Step 4
 Multiple indicator RI-CLPM, 5 waves with 3 indicators for each variable at each wave (30 observed variables). Fitting a model with constraints to ensure strong factorial invariance, with the RI-CLPM at the latent level.

```{r RICLPM-MI5, eval = F, warning = F}
RICLPM5.ext3 <- '
  
  #####################
  # MEASUREMENT MODEL #
  #####################
  
  # Factor models for x at 5 waves (constrained). 
  FX1 =~ a*x11 + b*x12 + c*x13
  FX2 =~ a*x21 + b*x22 + c*x23
  FX3 =~ a*x31 + b*x32 + c*x33
  FX4 =~ a*x41 + b*x42 + c*x43
  FX5 =~ a*x51 + b*x52 + c*x53
  
  # Factor models for y at 5 waves (constrained).
  FY1 =~ d*y11 + e*y12 + f*y13
  FY2 =~ d*y21 + e*y22 + f*y23
  FY3 =~ d*y31 + e*y32 + f*y33
  FY4 =~ d*y41 + e*y42 + f*y43
  FY5 =~ d*y51 + e*y52 + f*y53
  
  # Constrained intercepts over time (this is necessary for strong factorial 
  # invariance; without these contraints we have week factorial invariance). 
  
  x11 + x21 + x31 + x41 + x51 ~ g*1
  x12 + x22 + x32 + x42 + x52 ~ h*1
  x13 + x23 + x33 + x43 + x53 ~ i*1

  y11 + y21 + y31 + y41 + y51 ~ j*1
  y12 + y22 + y32 + y42 + y52 ~ k*1
  y13 + y23 + y33 + y43 + y53 ~ l*1

  # Free latent means from t = 2 onward (only do this in combination with the 
  # constraints on the intercepts; without these, this would not be specified).
  FX2 + FX3 + FX4 + FX5 + FY2 + FY3 + FY4 + FY5 ~ 1

  ################
  # BETWEEN PART #
  ################
  
  # Create between factors (random intercepts). 
  RIX =~ 1*FX1 + 1*FX2 + 1*FX3 + 1*FX4 + 1*FX5
  RIY =~ 1*FY1 + 1*FY2 + 1*FY3 + 1*FY4 + 1*FY5

  # Set the residual variances of all FX and FY variables to 0. 
  FX1 ~~ 0*FX1
  FX2 ~~ 0*FX2
  FX3 ~~ 0*FX3
  FX4 ~~ 0*FX4
  FX5 ~~ 0*FX5

  FY1 ~~ 0*FY1
  FY2 ~~ 0*FY2
  FY3 ~~ 0*FY3
  FY4 ~~ 0*FY4
  FY5 ~~ 0*FY5

  ###############
  # WITHIN PART #
  ###############
 
  # Create the within-part.
  WFX1 =~ 1*FX1
  WFX2 =~ 1*FX2
  WFX3 =~ 1*FX3
  WFX4 =~ 1*FX4
  WFX5 =~ 1*FX5
  
  WFY1 =~ 1*FY1
  WFY2 =~ 1*FY2
  WFY3 =~ 1*FY3
  WFY4 =~ 1*FY4
  WFY5 =~ 1*FY5
  
  # Specify the lagged effects between the within-person centered latent variables.
  WFX2 + WFY2 ~ WFX1 + WFY1
  WFX3 + WFY3 ~ WFX2 + WFY2
  WFX4 + WFY4 ~ WFX3 + WFY3
  WFX5 + WFY5 ~ WFX4 + WFY4
  
  # Estimate the correlations within the same wave.
  WFX2 ~~ WFY2
  WFX3 ~~ WFY3 
  WFX4 ~~ WFY4
  WFX5 ~~ WFY5
  
  ##########################
  # ADDITIONAL CONSTRAINTS #
  ##########################
  
  # Set correlations between the between-factors (random intercepts) and within-
  # factors at wave 1 at 0. 
  RIX + RIY ~~ 0*WFX1 + 0*WFY1

'
RICLPM5.ext3.fit <- cfa(RICLPM5.ext3, data = datMI, missing = 'ML')

summary(RICLPM5.ext3.fit, standardized = T)
```

---

## References

















