---
title: "RI-CLPM & Extensions using Lavaan"
output: 
  html_document:
    toc:  true
    toc_float: true
    code_folding: show
bibliography: references.bib
---

---

## Setup
Below you can find the code for installing and loading the required package `Lavaan` [@lavaan], as well as for reading in the data for the Random Intercept Cross-Lagged Panel Model (RI-CLPM) and its 3 extensions. You can specify the path to the data yourself, or through a menu by using the `file.choose()`-function. 

```{r setup, results = F, message = F}
# If necessary, install the 'Lavaan' package. 
# install.packages('lavaan', dependencies = T)

# Load the required packages. 
require(lavaan) 

# Load in the data. 
## Traditional RI-CLPM
dat <- read.table("C:/Users/5879167/surfdrive/Research/RI-CLPM/RI-CLPM/RICLPM.dat",
                  col.names = c("x1", "y1",
                                "x2", "y2", 
                                "x3", "y3",
                                "x4", "y4")) 
head(dat)

# Extension 1
datZ <- read.table("C:/Users/5879167/surfdrive/Research/RI-CLPM/RI-CLPMZ/RICLPMZ.dat",
                   col.names = c("x1", "x2", "x3", "x4",
                                "y1", "y2", "y3", "y4", 
                                "z1", "z2"))
head(datZ)

## Extension 2
datMG <- read.table("C:/Users/5879167/surfdrive/Research/RI-CLPM/MG RI-CLPM/MGRICLPM.dat",
                    col.names = c("x1", "x2", "x3", "x4",
                                "y1", "y2", "y3", "y4", 
                                "G"))
head(datMG)

# Extension 3
datMI <- read.table("C:/Users/5879167/surfdrive/Research/RI-CLPM/MI RI-CLPM/MIRICLPM.dat",
                   col.names = c("x11", "x12", "x13",
                                "x21", "x22", "x23",
                                "x31", "x32", "x33",
                                "x41", "x42", "x43",
                                
                                "y11", "y12", "y13",
                                "y21", "y22", "y23",
                                "y31", "y32", "y33", 
                                "y41", "y42", "y43"))
head(datMI)

```

---

## The RI-CLPM {.tabset .tabset-fade}

Use the tabs below to navigate to the model specification of either the basic model, or the RI-CLPM with constraints over time. 

### The basic model
The code for specifying the basic RI-CLPM is given below. 

```{r RICLPM.inp, results = F}
RICLPM <- '
  # Create two individual factors (random intercepts)
  RIx =~ 1*x1 + 1*x2 + 1*x3 + 1*x4
  RIy =~ 1*y1 + 1*y2 + 1*y3 + 1*y4
  
  # Create within-person centered variables
  wx1 =~ 1*x1
  wx2 =~ 1*x2
  wx3 =~ 1*x3 
  wx4 =~ 1*x4

  wy1 =~ 1*y1
  wy2 =~ 1*y2
  wy3 =~ 1*y3
  wy4 =~ 1*y4

  # Estimate the lagged effects between the within-
  # person centered variables.
  wx2 + wy2 ~ wx1 + wy1
  wx3 + wy3 ~ wx2 + wy2
  wx4 + wy4 ~ wx3 + wy3

  # Estimate the covariances between the residuals of 
  # the within-person centered variables (the innovations).
  wx2 ~~ wy2
  wx3 ~~ wy3
  wx4 ~~ wy4  

  # Estimate the covariance between the within-person 
  # centered variables at the first wave. 
  wx1 ~~ wy1 # Covariance
  
  # Estimate the variance and covariance of the individual 
  # factors. 
  RIx ~~ RIx
  RIy ~~ RIy
  RIx ~~ RIy

  # Estimate the (residual) variance of the within-person
  # centered variables.
  wx1 ~~ wx1 # Variance
  wy1 ~~ wy1 # Variance
  wx2 ~~ wx2 # Residual variance
  wy2 ~~ wy2 # Residual variance
  wx3 ~~ wx3 # Residual variance
  wy3 ~~ wy3 # Residual variance
  wx4 ~~ wx4 # Residual variance
  wy4 ~~ wy4 # Residual variance
'

RICLPM.fit <- lavaan(RICLPM, data = dat, missing = 'ML',
                     meanstructure = T, int.ov.free = T) 

summary(RICLPM.fit, standardized = T)
```

---

### Constraints over time
Below we specify a RI-CLPM with: 

1) fixed auto-regressive and cross-lagged relations over time, 
2) time-invariant (residual) (co-)variances in the within-person part, and 
3) constrained grand means over time. 

Please note that fixing parameters in Lavaan is done through **premultiplication**. It means that we have to add the number that we want to fix the parameter to and an asterisk, to the parameter in the model specification. For example, if we want to fix the factor loading of item `x1` to factor `F` to 0, we need to add `0*` to `F =~ x1`, resulting in `F =~ 0*x1`. 

```{r RICLPM3.inp, results = F}
RICLPM3 <- '
  # Create two individual factors (random intercepts)
  RIx =~ 1*x1 + 1*x2 + 1*x3 + 1*x4
  RIy =~ 1*y1 + 1*y2 + 1*y3 + 1*y4

  # Create within-person centered variables
  wx1 =~ 1*x1
  wx2 =~ 1*x2
  wx3 =~ 1*x3 
  wx4 =~ 1*x4

  wy1 =~ 1*y1
  wy2 =~ 1*y2
  wy3 =~ 1*y3
  wy4 =~ 1*y4

  # Estimate the lagged effects between the within-
  # person centered variables. 
  # Constrain the autoregressive effects and cross-
  # lagged effects to be invariant over time. 
  wx2 ~ a*wx1 + b*wy1
  wy2 ~ c*wx1 + d*wy1
  wx3 ~ a*wx2 + b*wy2
  wy3 ~ c*wx2 + d*wy2
  wx4 ~ a*wx3 + b*wy3
  wy4 ~ c*wx3 + d*wy3 

  # Estimate the covariances between the residuals of 
  # the within-person centered variables (the 
  # innovations), and constrain them over time. 
  wx2 ~~ cov*wy2
  wx3 ~~ cov*wy3
  wx4 ~~ cov*wy4  

  # Estimate the covariance between the within-person 
  # centered variables at the first wave. 
  wx1 ~~ wy1 # Covariance
  
  # Estimate the variance and covariance of the 
  # individual factors. 
  RIx ~~ RIx
  RIy ~~ RIy
  RIx ~~ RIy

  # Estimate the grand means, constrained over time 
  # (through `mx*` and `my*`).
  x1 + x2 + x3 + x4 ~ mx*1
  y1 + y2 + y3 + y4 ~ my*1  

  # Estimate the (residual) variance of the within-
  # person centered variables. Residual variances 
  # constrained to be invariant over time (through 
  # `vx*` and `vy*`). 
  wx1 ~~ wx1 # Variance
  wy1 ~~ wy1 # Variance
  wx2 ~~ vx*wx2 # Residual variance
  wy2 ~~ vy*wy2 # Residual variance
  wx3 ~~ vx*wx3 # Residual variance
  wy3 ~~ vy*wy3 # Residual variance
  wx4 ~~ vx*wx4 # Residual variance
  wy4 ~~ vy*wy4 # Residual variance
'

RICLPM3.fit <- lavaan(RICLPM3, data = dat, missing = 'ML',
                      meanstructure = T, int.ov.free = T) 

summary(RICLPM3.fit, standardized = T)
```

---

## Ext. 1: time-invariant predictor {.tabset .tabset-fade}

Use the tabs below to navigate to the model specification of the RI-CLPM with either $z_{1}$ as a predictor of the observed variables, or with $z_{1}$ as a predictor of the random intercepts.  

### $z_{1}$ as predictor of observed variables
Below you can find the code for the RI-CLPM with 4 waves and a between-level predictor for the observed variables. 

```{r RI-CLPM_observed_on_z1.inp, results = F}
RICLPM.ext1 <- '
  # Create two individual factors (random intercepts)
  RIx =~ 1*x1 + 1*x2 + 1*x3 + 1*x4
  RIy =~ 1*y1 + 1*y2 + 1*y3 + 1*y4
  
  # Create within-person centered variables
  wx1 =~ 1*x1
  wx2 =~ 1*x2
  wx3 =~ 1*x3 
  wx4 =~ 1*x4

  wy1 =~ 1*y1
  wy2 =~ 1*y2
  wy3 =~ 1*y3
  wy4 =~ 1*y4

  # Regression of observed variables on z1 (restricted). 
  x1 + x2 + x3 + x4 + y1 + y2 + y3+ y4 ~ z1 # Unrestricted. 
  # x1 + x2 + x3 + x4 ~ s1*z1 # Constrained over time. 
  # y1 + y2 + y3 + y4 ~ s2*z1 # Constrained over time. 

  # Estimate the lagged effects between the within-
  # person centered variables. 
  wx2 + wy2 ~ wx1 + wy1
  wx3 + wy3 ~ wx2 + wy2
  wx4 + wy4 ~ wx3 + wy3

  # Estimate the covariance between the within-person 
  # centered variables at the first wave. 
  wx1 ~~ wy1

  # Estimate the covariances between the residuals of 
  # the within-person centered variables (the innovations). 
  wx2 ~~ wy2
  wx3 ~~ wy3
  wx4 ~~ wy4  

  # Estimate the variance and covariance of the individual 
  # factors. 
  RIx ~~ RIx
  RIy ~~ RIy
  RIx ~~ RIy

  # Estimate the (residual) variance of the within-person
  # centered variables.
  wx1 ~~ wx1 # Variance
  wy1 ~~ wy1 # Variance
  wx2 ~~ wx2 # Residual variance
  wy2 ~~ wy2 # Residual variance
  wx3 ~~ wx3 # Residual variance
  wy3 ~~ wy3 # Residual variance
  wx4 ~~ wx4 # Residual variance
  wy4 ~~ wy4 # Residual variance
'

RICLPM.ext1.fit <- lavaan(RICLPM.ext1, data = datZ, missing = 'ML',
                          meanstructure = T, int.ov.free = T) 

summary(RICLPM.ext1.fit, standardized = T)
```

---

### $z_{1}$ as predictor of random intercepts
Below you can find the code for the RI-CLPM with 4 waves and the random intercepts regressed on predictor $z_{1}$. 
```{r RI-CLPM_RI_on_z1.inp, results = F}
RICLPM1.ext1 <- '
  # Create two individual factors (random intercepts)
  RIx =~ 1*x1 + 1*x2 + 1*x3 + 1*x4
  RIy =~ 1*y1 + 1*y2 + 1*y3 + 1*y4
  
  # Create within-person centered variables
  wx1 =~ 1*x1
  wx2 =~ 1*x2
  wx3 =~ 1*x3 
  wx4 =~ 1*x4

  wy1 =~ 1*y1
  wy2 =~ 1*y2
  wy3 =~ 1*y3
  wy4 =~ 1*y4

  # Regression of random intercepts on z1.  
  RIx + RIy ~ z1

  # Estimate the lagged effects between the within-
  # person centered variables. 
  wx2 + wy2 ~ wx1 + wy1
  wx3 + wy3 ~ wx2 + wy2
  wx4 + wy4 ~ wx3 + wy3

  # Estimate the covariance between the within-person
  # centered variables at the first wave. 
  wx1 ~~ wy1

  # Estimate the covariances between the residuals of
  # the within-person centered variables (the innovations). 
  wx2 ~~ wy2
  wx3 ~~ wy3
  wx4 ~~ wy4  

  # Estimate the variance and covariance of the individual
  # factors. 
  RIx ~~ RIx
  RIy ~~ RIy
  RIx ~~ RIy

  # Estimate the (residual) variance of the within-person
  # centered variables.
  wx1 ~~ wx1 # Variance
  wy1 ~~ wy1 # Variance
  wx2 ~~ wx2 # Residual variance
  wy2 ~~ wy2 # Residual variance
  wx3 ~~ wx3 # Residual variance
  wy3 ~~ wy3 # Residual variance
  wx4 ~~ wx4 # Residual variance
  wy4 ~~ wy4 # Residual variance
'

RICLPM1.ext1.fit <- lavaan(RICLPM1.ext1, data = datZ, missing = 'ML', 
                           meanstructure = T, int.ov.free = T) 

summary(RICLPM1.ext1.fit, standardized = T)
```

---

## Ext. 2: multiple group {.tabset .tabset-fade}

Use the tabs below to navigate to the model specification of the basic multiple-group model, or the model with constrained lagged parameters. 

### The basic model
Below you can find the code for a multiple group RI-CLPM with 4 waves.

```{r Multiple_group_RI-CLPM1.inp, results = F}
RICLPM1.ext2 <- '
  # Create two individual factors (random intercepts)
  RIx =~ 1*x1 + 1*x2 + 1*x3 + 1*x4
  RIy =~ 1*y1 + 1*y2 + 1*y3 + 1*y4
  
  # Create within-person centered variables
  wx1 =~ 1*x1
  wx2 =~ 1*x2
  wx3 =~ 1*x3 
  wx4 =~ 1*x4

  wy1 =~ 1*y1
  wy2 =~ 1*y2
  wy3 =~ 1*y3
  wy4 =~ 1*y4

  # Estimate the lagged effects between the within-
  # person centered variables. 
  wx2 + wy2 ~ wx1 + wy1
  wx3 + wy3 ~ wx2 + wy2
  wx4 + wy4 ~ wx3 + wy3

  # Estimate the covariance between the within-
  # person components at the first wave. 
  wx1 ~~ wy1 

  # Estimate the covariances between the residuals of
  # the within-person centered variables (the innovations). 
  wx2 ~~ wy2
  wx3 ~~ wy3
  wx4 ~~ wy4  

  # Estimate the variance and covariance of the individual
  # factors. 
  RIx ~~ RIx
  RIy ~~ RIy
  RIx ~~ RIy 

  # Estimate the (residual) variance of the within-person
  # centered variables.
  wx1 ~~ wx1 # Variance
  wy1 ~~ wy1 # Variance
  wx2 ~~ wx2 # Residual variance
  wy2 ~~ wy2 # Residual variance
  wx3 ~~ wx3 # Residual variance
  wy3 ~~ wy3 # Residual variance
  wx4 ~~ wx4 # Residual variance
  wy4 ~~ wy4 # Residual variance
'

RICLPM1.ext2.fit <- lavaan(RICLPM1.ext2, data = datMG, missing = 'ML',
                           group = "G", 
                           meanstructure = T, int.ov.free = T)
summary(RICLPM1.ext2.fit)

```

---

### Constrained lagged-parameters
Below you can find the code for a multiple group RI-CLPM with 4 waves. The lagged-parameters are constrained to be equal over time.

```{r Multiple_group_RI-CLPM2.inp, results = F}
RICLPM2.ext2 <- '
  # Create two individual factors (random intercepts).
  RIx =~ 1*x1 + 1*x2 + 1*x3 + 1*x4
  RIy =~ 1*y1 + 1*y2 + 1*y3 + 1*y4
  
  # Create within-person centered variables.
  wx1 =~ 1*x1
  wx2 =~ 1*x2
  wx3 =~ 1*x3 
  wx4 =~ 1*x4

  wy1 =~ 1*y1
  wy2 =~ 1*y2
  wy3 =~ 1*y3
  wy4 =~ 1*y4

  # Estimate the lagged effects between the within-
  # person centered variables. Constrain the  
  # autoregressive effects across groups. 
  wx2 ~ c(a1, a1)*wx1 + c(b1, b1)*wy1
  wy2 ~ c(c1, c1)*wx1 + c(d1, d1)*wy1
  wx3 ~ c(a2, a2)*wx2 + c(b2, b2)*wy2
  wy3 ~ c(c2, c2)*wx2 + c(d2, d2)*wy2
  wx4 ~ c(a3, a3)*wx3 + c(b3, b3)*wy3
  wy4 ~ c(c3, c3)*wx3 + c(d3, d3)*wy3 

  # Estimate the covariance between the within-person
  # components at the first wave. 
  wx1 ~~ wy1 

 # Estimate the covariances between the residuals of
  # the within-person centered variables (the innovations). 
  wx2 ~~ wy2
  wx3 ~~ wy3
  wx4 ~~ wy4  

  # Estimate the variance and covariance of the individual
  # factors. 
  RIx ~~ RIx
  RIy ~~ RIy
  RIx ~~ RIy 

  # Estimate the (residual) variance of the within-person
  # centered variables.
  wx1 ~~ wx1 # Variance
  wy1 ~~ wy1 # Variance
  wx2 ~~ wx2 # Residual variance
  wy2 ~~ wy2 # Residual variance
  wx3 ~~ wx3 # Residual variance
  wy3 ~~ wy3 # Residual variance
  wx4 ~~ wx4 # Residual variance
  wy4 ~~ wy4 # Residual variance
'

RICLPM2.ext2.fit <- lavaan(RICLPM2.ext2, data = datMG, missing = 'ML',
                           group = "G", 
                           meanstructure = T, int.ov.free = T)
summary(RICLPM2.ext2.fit)

```

---

## Ext. 3: multiple indicator {.tabset .tabset-fade}

Use the tabs below to navigate to the model specification of a multiple indicator RI-CLPM with 4 waves and 3 indicators for each variable at each wave. The five steps correspond to models with

1. Initial model (Step 1).
2. Weak factorial invariance (Step 2).
3. Strong factorial invariance (Step 3).
4. Strong factorial invariance with factor loadings equal to the within-person factor loadings (Extra). 
5. The latent RI-CLPM (Step 4).

### Step 1
Multiple indicator RI-CLPM 4 waves with 3 indicators for each variable at each wave (24 observed variables), with random intercepts for each indicator separately.

```{r mi_ri-clpm_step_1.inp, results = F, warning = F}
RICLPM1.ext3 <- '
  
  ################
  # BETWEEN PART #
  ################
  
  # Create between factors (random intercepts) for each
  # indicator separately.
  RIX1 =~ 1*x11 + 1*x21 + 1*x31 + 1*x41
  RIX2 =~ 1*x12 + 1*x22 + 1*x32 + 1*x42
  RIX3 =~ 1*x13 + 1*x23 + 1*x33 + 1*x43
  
  RIY1 =~ 1*y11 + 1*y21 + 1*y31 + 1*y41
  RIY2 =~ 1*y12 + 1*y22 + 1*y32 + 1*y42
  RIY3 =~ 1*y13 + 1*y23 + 1*y33 + 1*y43
  
  ##################################
  # WITHIN PART: MEASUREMENT MODEL #
  ##################################
  
  # Factor models for x at 4 waves.
  WFX1 =~ x11 + x12 + x13
  WFX2 =~ x21 + x22 + x23
  WFX3 =~ x31 + x32 + x33
  WFX4 =~ x41 + x42 + x43
  
  # Factor models for y at 4 waves.
  WFY1 =~ y11 + y12 + y13
  WFY2 =~ y21 + y22 + y23
  WFY3 =~ y31 + y32 + y33
  WFY4 =~ y41 + y42 + y43
  
  #########################
  # WITHIN PART: DYNAMICS #
  #########################
  
  # Specify the lagged effects between the within-person 
  # centered latent variables.
  WFX2 + WFY2 ~ WFX1 + WFY1
  WFX3 + WFY3 ~ WFX2 + WFY2
  WFX4 + WFY4 ~ WFX3 + WFY3
  
  # Estimate the correlations within the same wave.
  WFX1 ~~ WFY1
  WFX2 ~~ WFY2
  WFX3 ~~ WFY3 
  WFX4 ~~ WFY4
  
  ##########################
  # ADDITIONAL CONSTRAINTS #
  ##########################
  
  # Constrain covariance of the the between factors and
  # exogenous within factors to 0. 
  RIX1 + RIX2 + RIX3 + RIY1 + RIY2 + RIY3 ~~ 0*WFY1 + 0*WFX1

'
RICLPM1.ext3.fit <- cfa(RICLPM1.ext3, data = datMI, missing = 'ML')

summary(RICLPM1.ext3.fit, standardized = T)
```

---

### Step 2
Multiple indicator RI-CLPM 4 waves with 3 indicators for each variable at each wave (24 observed variables). Fitting a model with constraints to ensure weak factorial invariance, and with a random intercept for each indicator separately.

```{r MI_RI-CLPM_step_2.inp, results = F, warning = F}
RICLPM2.ext3 <- '
  
  ################
  # BETWEEN PART #
  ################
  
  # Create between factors (random intercepts) for each
  # indicator separately.
  RIX1 =~ 1*x11 + 1*x21 + 1*x31 + 1*x41
  RIX2 =~ 1*x12 + 1*x22 + 1*x32 + 1*x42
  RIX3 =~ 1*x13 + 1*x23 + 1*x33 + 1*x43
  
  RIY1 =~ 1*y11 + 1*y21 + 1*y31 + 1*y41
  RIY2 =~ 1*y12 + 1*y22 + 1*y32 + 1*y42
  RIY3 =~ 1*y13 + 1*y23 + 1*y33 + 1*y43
  
  ##################################
  # WITHIN PART: MEASUREMENT MODEL #
  ##################################
  
  # Factor models for x at 4 waves: constrained factor 
  # loadings over time.
  WFX1 =~ a*x11 + b*x12 + c*x13
  WFX2 =~ a*x21 + b*x22 + c*x23
  WFX3 =~ a*x31 + b*x32 + c*x33
  WFX4 =~ a*x41 + b*x42 + c*x43
  
  # Factor models for y at 4 waves.
  WFY1 =~ d*y11 + e*y12 + f*y13
  WFY2 =~ d*y21 + e*y22 + f*y23
  WFY3 =~ d*y31 + e*y32 + f*y33
  WFY4 =~ d*y41 + e*y42 + f*y43
  
  #########################
  # WITHIN PART: DYNAMICS #
  #########################
 
  # Specify the lagged effects between the within-person 
  # centered latent variables.
  WFX2 + WFY2 ~ WFX1 + WFY1
  WFX3 + WFY3 ~ WFX2 + WFY2
  WFX4 + WFY4 ~ WFX3 + WFY3
  
  # Estimate the correlations within the same wave.
  WFX1 ~~ WFY1
  WFX2 ~~ WFY2
  WFX3 ~~ WFY3 
  WFX4 ~~ WFY4
  
  ##########################
  # ADDITIONAL CONSTRAINTS #
  ##########################
  
  # Constrain covariance of the the between factors and 
  # exogenous within factors to 0. 
  RIX1 + RIX2 + RIX3 + RIY1 + RIY2 + RIY3 ~~ 0*WFY1 + 0*WFX1

'
RICLPM2.ext3.fit <- cfa(RICLPM2.ext3, data = datMI, missing = 'ML')

summary(RICLPM2.ext3.fit, standardized = T)
```

---

### Step 3
Multiple indicator RI-CLPM 4 waves with 3 indicators for each variable at each wave (24 observed variables). Fitting a model with constraints to ensure strong factorial invariance, with a random intercept for each indicator separately.

```{r MI_RI-CLPM_step_3.inp, results = F, warning = F}
RICLPM3.ext3 <- '
  ################
  # BETWEEN PART #
  ################
  
  # Create between factors (random intercepts) for
  # each indicator separately.
  RIX1 =~ 1*x11 + 1*x21 + 1*x31 + 1*x41
  RIX2 =~ 1*x12 + 1*x22 + 1*x32 + 1*x42
  RIX3 =~ 1*x13 + 1*x23 + 1*x33 + 1*x43
  
  RIY1 =~ 1*y11 + 1*y21 + 1*y31 + 1*y41
  RIY2 =~ 1*y12 + 1*y22 + 1*y32 + 1*y42
  RIY3 =~ 1*y13 + 1*y23 + 1*y33 + 1*y43
  
  ##################################
  # WITHIN PART: MEASUREMENT MODEL #
  ##################################
  
  # Factor models for x at 4 waves: constrained factor 
  # loadings over time.
  WFX1 =~ a*x11 + b*x12 + c*x13
  WFX2 =~ a*x21 + b*x22 + c*x23
  WFX3 =~ a*x31 + b*x32 + c*x33
  WFX4 =~ a*x41 + b*x42 + c*x43
  
  # Factor models for y at 4 waves.
  WFY1 =~ d*y11 + e*y12 + f*y13
  WFY2 =~ d*y21 + e*y22 + f*y23
  WFY3 =~ d*y31 + e*y32 + f*y33
  WFY4 =~ d*y41 + e*y42 + f*y43
  
  # Constrained intercepts over time (this is necessary
  # for strong factorial invariance; without these 
  # contraints we have week factorial invariance). 
  
  x11 + x21 + x31 + x41 ~ g*1
  x12 + x22 + x32 + x42 ~ h*1
  x13 + x23 + x33 + x43 ~ i*1

  y11 + y21 + y31 + y41 ~ j*1
  y12 + y22 + y32 + y42 ~ k*1
  y13 + y23 + y33 + y43 ~ l*1
  
  # Free latent means from t = 2 onward (only do this in 
  # combination with the constraints on the intercepts; 
  # without these, this would not be specified).
  
  WFX2 + WFX3 + WFX4 + WFY2 + WFY3 + WFY4 ~ 1
  
  #########################
  # WITHIN PART: DYNAMICS #
  #########################
  
  # Specify the lagged effects between the within-person
  # centered latent variables.
  WFX2 + WFY2 ~ WFX1 + WFY1
  WFX3 + WFY3 ~ WFX2 + WFY2
  WFX4 + WFY4 ~ WFX3 + WFY3
  
  # Estimate the correlations within the same wave.
  WFX1 ~~ WFY1
  WFX2 ~~ WFY2
  WFX3 ~~ WFY3 
  WFX4 ~~ WFY4
  
  # Constrain covariance of the the between factors and
  # exogenous within factors to 0. 
  RIX1 + RIX2 + RIX3 + RIY1 + RIY2 + RIY3 ~~ 0*WFY1 + 0*WFX1

'
RICLPM3.ext3.fit <- cfa(RICLPM3.ext3, data = datMI, missing = 'ML')

summary(RICLPM3.ext3.fit, standardized = T)
```

---

### Extra
Multiple indicator RI-CLPM 4 waves with 3 indicators for each variable at each wave (24 observed variables). Fitting a model with constraints to ensure strong factorial invariance, with a random intercept for each indicator separately, for which a factor model is specified, with factor loadings equal to the within-person factor loadings.

```{r MI_RI-CLPM_step_4.inp, results = F, warning = F}
RICLPM4.ext3 <- '
  ################
  # BETWEEN PART #
  ################
  
  # Create between factors (random intercepts) for each
  # indicator separately.
  RIX1 =~ 1*x11 + 1*x21 + 1*x31 + 1*x41
  RIX2 =~ 1*x12 + 1*x22 + 1*x32 + 1*x42
  RIX3 =~ 1*x13 + 1*x23 + 1*x33 + 1*x43
  
  RIY1 =~ 1*y11 + 1*y21 + 1*y31 + 1*y41
  RIY2 =~ 1*y12 + 1*y22 + 1*y32 + 1*y42
  RIY3 =~ 1*y13 + 1*y23 + 1*y33 + 1*y43
  
  # Create a single random intercept for al x variables, 
  # and another for all y variables and constrain the factor 
  # loadings to be identical to the within-person factor 
  # loadings. 
  
  RIX =~ a*RIX1 + b*RIX2 + c*RIX3
  RIY =~ d*RIY1 + e*RIY2 + f*RIY3
  
  # Constrain measurement error variances of the second order
  # factor model to 0. 
  RIX1 ~~ 0*RIX1
  RIX2 ~~ 0*RIX2
  RIX3 ~~ 0*RIX3

  RIY1 ~~ 0*RIY1
  RIY2 ~~ 0*RIY2
  RIY3 ~~ 0*RIY3
  
  ##################################
  # WITHIN PART: MEASUREMENT MODEL #
  ##################################
  
  # Due to having a random intercept per indicator, the 
  # measurement model here is only on the within part of 
  # the measurements. 
  
  # Factor models for x at 4 waves: constrained factor 
  # loadings over time.
  WFX1 =~ a*x11 + b*x12 + c*x13
  WFX2 =~ a*x21 + b*x22 + c*x23
  WFX3 =~ a*x31 + b*x32 + c*x33
  WFX4 =~ a*x41 + b*x42 + c*x43
  
  # Factor models for y at 4 waves.
  WFY1 =~ d*y11 + e*y12 + f*y13
  WFY2 =~ d*y21 + e*y22 + f*y23
  WFY3 =~ d*y31 + e*y32 + f*y33
  WFY4 =~ d*y41 + e*y42 + f*y43
  
  # Constrained intercepts over time (this is necessary for 
  # strong factorial invariance; without these contraints 
  # we have week factorial invariance). 
  
  x11 + x21 + x31 + x41 ~ g*1
  x12 + x22 + x32 + x42 ~ h*1
  x13 + x23 + x33 + x43 ~ i*1

  y11 + y21 + y31 + y41 ~ j*1
  y12 + y22 + y32 + y42 ~ k*1
  y13 + y23 + y33 + y43 ~ l*1
  
  # Free latent means from t = 2 onward (only do this in 
  # combination with the constraints on the intercepts; 
  # without these, this would not be specified).
  
  WFX2 + WFX3 + WFX4 + WFY2 + WFY3 + WFY4 ~ 1
  
  #########################
  # WITHIN PART: DYNAMICS #
  #########################
  
  # Specify the lagged effects between the within-person 
  # centered latent variables.
  WFX2 + WFY2 ~ WFX1 + WFY1
  WFX3 + WFY3 ~ WFX2 + WFY2
  WFX4 + WFY4 ~ WFX3 + WFY3
  
  # Estimate the correlations within the same wave.
  WFX2 ~~ WFY2
  WFX3 ~~ WFY3 
  WFX4 ~~ WFY4
  
  ##########################
  # ADDITIONAL CONSTRAINTS #
  ##########################
  
  # Set correlations between the between-factors (random 
  # intercepts) and within-factors at wave 1 at 0. 
  RIX + RIY ~~ 0*WFX1 + 0*WFY1

'
RICLPM4.ext3.fit <- cfa(RICLPM4.ext3, data = datMI, missing = 'ML')

summary(RICLPM4.ext3.fit, standardized = T)
```

---

### Step 4
Multiple indicator RI-CLPM with 4 waves and 3 indicators for each variable at each wave (24 observed variables). Fitting a model with constraints to ensure strong factorial invariance, with the RI-CLPM at the latent level. 

```{r MI_RI-CLPM_step_5.inp, results = F, warning = F}
RICLPM5.ext3 <- '
  #####################
  # MEASUREMENT MODEL #
  #####################
  
  # Factor models for x at 4 waves: constrained factor 
  # loadings over time.
  FX1 =~ a*x11 + b*x12 + c*x13
  FX2 =~ a*x21 + b*x22 + c*x23
  FX3 =~ a*x31 + b*x32 + c*x33
  FX4 =~ a*x41 + b*x42 + c*x43
  
  # Factor models for y at 4 waves.
  FY1 =~ d*y11 + e*y12 + f*y13
  FY2 =~ d*y21 + e*y22 + f*y23
  FY3 =~ d*y31 + e*y32 + f*y33
  FY4 =~ d*y41 + e*y42 + f*y43
  
  # Constrained intercepts over time (this is necessary 
  # for strong factorial invariance; without these contraints
  # we have week factorial invariance). 
  
  x11 + x21 + x31 + x41 ~ g*1
  x12 + x22 + x32 + x42 ~ h*1
  x13 + x23 + x33 + x43 ~ i*1

  y11 + y21 + y31 + y41 ~ j*1
  y12 + y22 + y32 + y42 ~ k*1
  y13 + y23 + y33 + y43 ~ l*1

  # Free latent means from t = 2 onward (only do this in 
  # combination with the constraints on the intercepts; without
  # these, this would not be specified).
  FX2 + FX3 + FX4 + FY2 + FY3 + FY4 ~ 1

  ################
  # BETWEEN PART #
  ################
  
  # Create between factors (random intercepts). 
  RIX =~ 1*FX1 + 1*FX2 + 1*FX3 + 1*FX4
  RIY =~ 1*FY1 + 1*FY2 + 1*FY3 + 1*FY4

  # Set the residual variances of all FX and FY variables to 0. 
  FX1 ~~ 0*FX1
  FX2 ~~ 0*FX2
  FX3 ~~ 0*FX3
  FX4 ~~ 0*FX4

  FY1 ~~ 0*FY1
  FY2 ~~ 0*FY2
  FY3 ~~ 0*FY3
  FY4 ~~ 0*FY4

  ###############
  # WITHIN PART #
  ###############
 
  # Create the within-part.
  WFX1 =~ 1*FX1
  WFX2 =~ 1*FX2
  WFX3 =~ 1*FX3
  WFX4 =~ 1*FX4
  
  WFY1 =~ 1*FY1
  WFY2 =~ 1*FY2
  WFY3 =~ 1*FY3
  WFY4 =~ 1*FY4
  
  # Specify the lagged effects between the within-person 
  # centered latent variables.
  WFX2 + WFY2 ~ WFX1 + WFY1
  WFX3 + WFY3 ~ WFX2 + WFY2
  WFX4 + WFY4 ~ WFX3 + WFY3
  
  # Estimate the correlations within the same wave.
  WFX2 ~~ WFY2
  WFX3 ~~ WFY3 
  WFX4 ~~ WFY4
  
  ##########################
  # ADDITIONAL CONSTRAINTS #
  ##########################
  
  # Set correlations between the between-factors (random 
  # intercepts) and within-factors at wave 1 at 0. 
  RIX + RIY ~~ 0*WFX1 + 0*WFY1

'
RICLPM5.ext3.fit <- cfa(RICLPM5.ext3, data = datMI, missing = 'ML')

summary(RICLPM5.ext3.fit, standardized = T)
```

---

## References

















